<!-- # Child Rmarkdown Document for Cluster Annotation -->
<!-- Parent document must have a variable "in_rna" containing metadata-injected H5 files by sample --> 
<a id="rna_seq_cluster_annotation_top"></a>

```{r, include=FALSE} 
orig_workspace <- ls()

scrna_seq_sample_module_version <- "1.0.2" # 20211201
stm("Starting scRNA Cluster Annotation Module")
```

### Contents 
#### [QC Seurat Object](#seurat_qc)
  - [Principal Component Plot (Elbow Plot)](#elbow_plot)
  - [UMAP-Treatments](#UMAP)
  - [Marker Gene Table](#mgt) 
  - [UMAP with Cell Annotations](#cell_type_annotation)
<details style="color: lightgray;">  
  <summary>Expand Code</summary> 


```{r rna_markergenes}
stm("Reading in Marker Genes")

if (length(grep("https",in_key)) > 0) {
    ss <- read_sheet(in_key)
    } else if (length(grep(".xlsx",in_key)) > 0 ){
        ss <- import_list(in_key)
    } else {
        ss <- read.csv(in_key)
}

metatable <- ss$MetaTable_expectedCell
if ("Final list" %in% colnames(metatable)) {
    metatable <- metatable %>% filter(metatable$"Final list" == 1)
}
samples <- unique(metatable$Sample)
treatments <- metatable$Treatment      
              
# Format Marker Genes
MarkerGenes <- ss$MarkerGenes
colnames(MarkerGenes) <- c('cell_type','marker_gene')
markers <- MarkerGenes$'marker_gene'

# read in gene name table
geneTable <- read.csv(paste0(refdir, "geneAnnotationTable.csv"), header = T, row.names = 1)
``` 
  
```{r }
# scCustermize code updated by developer, https://github.com/samuel-marsh/scCustomize/issues/27
# Clustered_DotPlot_relabel function from developer

Clustered_DotPlot_relabel <- function(
  seurat_object,
  features,
  new_row_labels,
  colors_use_exp = viridis_plasma_dark_high,
  exp_color_min = -2,
  exp_color_middle = NULL,
  exp_color_max = 2,
  print_exp_quantiles = FALSE,
  colors_use_idents = NULL,
  x_lab_rotate = TRUE,
  k = 1,
  row_km_repeats = 1000,
  column_km_repeats = 1000,
  row_label_size = 8,
  raster = FALSE,
  plot_km_elbow = TRUE,
  elbow_kmax = NULL,
  assay = NULL,
  group.by = NULL,
  idents = NULL,
  show_parent_dend_line = TRUE,
  ggplot_default_colors = FALSE,
  seed = 123
) {
  # Check for packages
  ComplexHeatmap_check <- PackageCheck("ComplexHeatmap", error = FALSE)
  if (!ComplexHeatmap_check[1]) {
    stop(
      "Please install the ComplexHeatmap package to use Clustered_DotPlot",
      "\nThis can be accomplished with the following commands: ",
      "\n----------------------------------------",
      "\ninstall.packages('BiocManager')",
      "\nBiocManager::install('ComplexHeatmap')",
      "\n----------------------------------------",
      call. = FALSE
    )
  }
  
  # Check Seurat
  scCustomize:::Is_Seurat(seurat_object = seurat_object)
  
  # Check unique features
  features_unique <- unique(x = features)
  
  if (length(x = features_unique) != length(x = features)) {
    warning("Feature list contains duplicates, making unique.")
  }
  
  # Check exp min/max set correctly
  if (!exp_color_min < exp_color_max) {
    stop("The value for 'exp_color_min': ", exp_color_min, ", must be less than the value for 'exp_color_max': ", exp_color_max, ".")
  }
  
  # Get DotPlot data
  seurat_plot <- DotPlot(object = seurat_object, features = features_unique, assay = assay, group.by = group.by, scale = TRUE, idents = idents, col.min = NULL, col.max = NULL)
  
  data <- seurat_plot$data
  
  # Get expression data
  exp_mat <- data %>%
    dplyr::select(-pct.exp, -avg.exp) %>%
    pivot_wider(names_from = id, values_from = avg.exp.scaled) %>%
    as.data.frame()
  
  row.names(x = exp_mat) <- exp_mat$features.plot
  
  # Check NAs if idents
  if (!is.null(x = idents)) {
    # Find NA features and print warning
    excluded_features <- exp_mat[rowSums(is.na(x = exp_mat)) > 0,] %>%
      rownames()
    warning("The following features were removed as there is no scaled expression present in subset (`idents`) of object provided: ", glue_collapse_scCustom(input_string = excluded_features, and = TRUE), ".")
    
    # Extract good features
    good_features <- rownames(exp_mat)
    
    # Remove rows with NAs
    exp_mat <- exp_mat %>%
      filter(features.plot %in% good_features)
  }
  
  exp_mat <- exp_mat[,-1] %>%
    as.matrix()
  
  # Get percent expressed data
  percent_mat <- data %>%
    dplyr::select(-avg.exp, -avg.exp.scaled) %>%
    pivot_wider(names_from = id, values_from = pct.exp) %>%
    as.data.frame()
  
  row.names(x = percent_mat) <- percent_mat$features.plot
  
  # Subset dataframe for NAs if idents so that exp_mat and percent_mat match
  if (!is.null(x = idents)) {
    percent_mat <- percent_mat %>%
      filter(features.plot %in% good_features)
  }
  
  percent_mat <- percent_mat[,-1] %>%
    as.matrix()
  
  # print quantiles
  if (print_exp_quantiles) {
    message("Quantiles of gene expression data are:")
    print(quantile(exp_mat, c(0.1, 0.5, 0.9, 0.99)))
  }
  
  # set assay (if null set to active assay)
  assay <- assay %||% DefaultAssay(object = seurat_object)
  
  # Set default color palette based on number of levels being plotted
  if (is.null(x = group.by)) {
    group_by_length <- length(x = unique(x = seurat_object@active.ident))
  } else {
    group_by_length <- length(x = unique(x = seurat_object@meta.data[[group.by]]))
  }
  
  # Check colors use vs. ggplot2 color scale
  if (!is.null(x = colors_use_idents) && ggplot_default_colors) {
    stop("Cannot provide both custom palette to `colors_use` and specify `ggplot_default_colors = TRUE`.")
  }
  if (is.null(x = colors_use_idents)) {
    # set default plot colors
    if (is.null(x = colors_use_idents)) {
      colors_use_idents <- scCustomize_Palette(num_groups = group_by_length, ggplot_default_colors = ggplot_default_colors, color_seed = color_seed)
    }
  }
  
  # Pull Annotation and change colors to ComplexHeatmap compatible format
  Identity <- colnames(exp_mat)
  
  identity_colors <- DiscretePalette_scCustomize(num_colors = length(Identity), palette = "polychrome", shuffle_pal = F)
  names(identity_colors) <- Identity
  identity_colors_list <- list(Identity = identity_colors)
  
  # Create identity annotation
  column_ha <- ComplexHeatmap::HeatmapAnnotation(Identity = Identity,
                                                 col =  identity_colors_list,
                                                 na_col = "grey",
                                                 name = "Identity"
  )
  
  # Set middle of color scale if not specified
  if (is.null(x = exp_color_middle)) {
    exp_color_middle <- scCustomize:::Middle_Number(min = exp_color_min, max = exp_color_max)
  }
  
  palette_length <- length(colors_use_exp)
  palette_middle <- scCustomize:::Middle_Number(min = 0, max = palette_length)
  
  # Create palette
  col_fun = colorRamp2(c(exp_color_min, exp_color_middle, exp_color_max), colors_use_exp[c(1,palette_middle, palette_length)])
  
  # Calculate and plot Elbow
  if (plot_km_elbow) {
    # if elbow_kmax not NULL check it is usable
    if (!is.null(x = elbow_kmax) && elbow_kmax > (nrow(x = exp_mat) - 1)) {
      elbow_kmax <- nrow(x = exp_mat) - 1
      warning("The value provided for 'elbow_kmax' is too large.  Changing to (length(x = features)-1): ", elbow_kmax)
    }
    
    # if elbow_kmax is NULL set value based on input feature list
    if (is.null(x = elbow_kmax)) {
      # set to (length(x = features)-1) if less than 21 features OR to 20 if greater than 21 features
      if (nrow(x = exp_mat) > 21) {
        elbow_kmax <- 20
      } else {
        elbow_kmax <- nrow(x = exp_mat) - 1
      }
    }
    
    km_elbow_plot <- scCustomize:::kMeans_Elbow(data = exp_mat, k_max = elbow_kmax)
  }
  
  # prep heatmap
  if (raster) {
    layer_fun = function(j, i, x, y, w, h, fill) {
      grid.rect(x = x, y = y, width = w, height = h,
                gp = gpar(col = NA, fill = NA))
      grid.circle(x=x,y=y,r= sqrt(ComplexHeatmap::pindex(percent_mat, i, j)/100)  * unit(2, "mm"),
                  gp = gpar(fill = col_fun(ComplexHeatmap::pindex(exp_mat, i, j)), col = NA))
    }
  } else {
    cell_fun = function(j, i, x, y, w, h, fill) {
      grid.rect(x = x, y = y, width = w, height = h,
                gp = gpar(col = NA, fill = NA))
      grid.circle(x=x,y=y,r= sqrt(percent_mat[i, j]/100) * unit(2, "mm"),
                  gp = gpar(fill = col_fun(exp_mat[i, j]), col = NA))
    }
  }
  
  # Create legend for point size
  lgd_list = list(
    ComplexHeatmap::Legend(labels = c(0.25,0.5,0.75,1), title = "Percent Expressing",
                           graphics = list(
                             function(x, y, w, h) grid.circle(x = x, y = y, r = sqrt(0.25) * unit(2, "mm"),
                                                              gp = gpar(fill = "black")),
                             function(x, y, w, h) grid.circle(x = x, y = y, r = sqrt(0.5) * unit(2, "mm"),
                                                              gp = gpar(fill = "black")),
                             function(x, y, w, h) grid.circle(x = x, y = y, r = sqrt(0.75) * unit(2, "mm"),
                                                              gp = gpar(fill = "black")),
                             function(x, y, w, h) grid.circle(x = x, y = y, r = 1 * unit(2, "mm"),
                                                              gp = gpar(fill = "black")))
    )
  )
  
  # Set x label roration
  if (is.numeric(x = x_lab_rotate)) {
    x_lab_rotate <- x_lab_rotate
  } else if (isTRUE(x = x_lab_rotate)) {
    x_lab_rotate <- 45
  } else {
    x_lab_rotate <- 0
  }
  
  # Create Plot
  set.seed(seed = seed)
  if (raster) {
    cluster_dot_plot <- ComplexHeatmap::Heatmap(exp_mat,
                                                heatmap_legend_param=list(title="Expression"),
                                                col=col_fun,
                                                rect_gp = gpar(type = "none"),
                                                layer_fun = layer_fun,
                                                row_names_gp = gpar(fontsize = row_label_size),
                                                row_km = k,
                                                row_km_repeats = row_km_repeats,
                                                border = "black",
                                                top_annotation = column_ha,
                                                column_km_repeats = column_km_repeats,
                                                show_parent_dend_line = show_parent_dend_line,
                                                column_names_rot = x_lab_rotate)
  } else {
    cluster_dot_plot <- ComplexHeatmap::Heatmap(exp_mat,
                                                heatmap_legend_param=list(title="Expression"),
                                                col=col_fun,
                                                rect_gp = gpar(type = "none"),
                                                cell_fun = cell_fun,
                                                row_names_gp = gpar(fontsize = row_label_size),
                                                row_km = k,
                                                row_km_repeats = row_km_repeats,
                                                border = "black",
                                                top_annotation = column_ha,
                                                column_km_repeats = column_km_repeats,
                                                show_parent_dend_line = show_parent_dend_line,
                                                column_names_rot = x_lab_rotate)
  }
  
  # Add pt.size legend & return plots
  if (plot_km_elbow) {
    return(list(km_elbow_plot, ComplexHeatmap::draw(cluster_dot_plot, annotation_legend_list = lgd_list)))
  }
  return(ComplexHeatmap::draw(cluster_dot_plot + rowAnnotation(rn= anno_text(new_row_labels)), annotation_legend_list = lgd_list))
}
```
  

```{r pre_processing}
### Pre-Processing
# Remove Ambient RNA by SoupX
stm('Preprocesing, Removing Ambient RNA by SoupX')
# Step 2: Pre-processing
# Remove ambient RNA by SoupX
setwd(cellrangers_dir)
data.10x = list()
for (sample in samples){
  filt.matrix <- Read10X_h5(paste0(cellrangers_dir, sample, "/outs/filtered_feature_bc_matrix.h5"), use.names = T)
  raw.matrix <- Read10X_h5(paste0(cellrangers_dir, sample, "/outs/raw_feature_bc_matrix.h5"), use.names = T)
  srat <- CreateSeuratObject(counts = filt.matrix)
  soup.channel <- SoupChannel(raw.matrix, filt.matrix)
  srat <- SCTransform(srat, verbose = F)
  srat <- RunPCA(srat, verbose = F)
  srat <- RunUMAP(srat, dims = 1:30, verbose = F)
  srat <- FindNeighbors(srat, dims = 1:30, verbose = F)
  srat <- FindClusters(srat, verbose = F)
  meta <- srat@meta.data
  umap <- srat@reductions$umap@cell.embeddings
  soup.channel <- setClusters(soup.channel, setNames(meta$seurat_clusters, rownames(meta)))
  soup.channel <- setDR(soup.channel, umap)
  soup.channel <- autoEstCont(soup.channel)
  data.10x[[sample]] <- adjustCounts(soup.channel, roundToInt = T)
}

# Create Seurat object after SoupX
scrna.list = list()
for (sample in samples) {
    scrna.list[[sample]] = CreateSeuratObject(counts = data.10x[[sample]], min.cells=3, project=sample)
}

# Remove raw data to save memory
rm(data.10x)
                        
# Read in doublet scores
for (sample in samples){
  doublet_scores <- scan(paste0(scrubletdir, sample, "_scrublet.score"))
  predicted_doublets <- scan(paste0(scrubletdir, sample, "_scrublet.logic"))
  ds <- as.data.frame(cbind(doublet_scores, predicted_doublets))
  ds$predicted_doublets <- as.logical(ds$predicted_doublets)
  rownames(ds) <- rownames(scrna.list[[sample]]@meta.data)
  scrna.list[[sample]] <- AddMetaData(scrna.list[[sample]], ds)
  scrna.list[[sample]] <- subset(scrna.list[[sample]], subset=predicted_doublets == FALSE)
}

# add sample name
for(i in 1:length(samples)){
  sample=samples[i]; treatment=treatments[i];
  scrna.list[[sample]]$treatment <- treatment
}

                                    
                                                                  
rm(ds)
rm(filt.matrix)
rm(meta)
rm(raw.matrix)
rm(soup.channel)
rm(srat)
rm(doublet_scores)
rm(predicted_doublets)

```

[Return to Contents](#rna_seq_sample_top) 

```{r QC}
rm(data.10x)
# Add percent.mt and percent.rb to cell level metadata
for (sample in samples) {
  scrna.list[[sample]][["percent.mito"]] <- PercentageFeatureSet(scrna.list[[sample]], pattern = "^MT:|MT-") 
  scrna.list[[sample]][["percent.ribo"]] <- PercentageFeatureSet(scrna.list[[sample]], pattern = "^RP[LS]")
}
# merge list of prefiltered Seurat
scrna.combined_prefilter <- Merge_Seurat_List(scrna.list,
                                              add.cell.ids = NULL,
                                              merge.data = TRUE,
                                              project = "orig.ident")
# metadata variable
metadata_prefilter <- scrna.combined_prefilter@meta.data
# Save Prefilter
saveRDS(scrna.combined_prefilter, paste0(outdir, "scrna.combined_prefilter.seurat.", projectName, ".rds"))
                   
# Filtered Cells with 3SD of mean nCount and nFeature, percent of mito
  
so_postfilter <- subset(x = scrna.combined_prefilter, 
                            subset= (nCount_RNA >= nUMI) & 
                           (nFeature_RNA >= nFeature_RNA ) &
                           (percent.mito <= percent_mito))

metadata_postfilter <- so_postfilter@meta.data

```
### QC Plots
<a  id="Mitochondrial_Percentage"></a> 
### Mitochondrial Percentage
```{r Mitochondrial Percentage Plot}
# mito
prefilter_qc_mito <- metadata_prefilter %>% ggplot(aes(x=nCount_RNA, y=nFeature_RNA, color=percent.mito)) + 
geom_point() + 
scale_colour_gradient(low = "gray90", high = "black") +
stat_smooth(method=lm) +
scale_x_log10() +
scale_y_log10() + 
theme_classic() +
geom_vline(xintercept = 500) +
geom_hline(yintercept = 300) +
facet_wrap(~orig.ident) + 
ggtitle("Prefilter")

postfilter_qc_mito <- metadata_postfilter %>% ggplot(aes(x=nCount_RNA, y=nFeature_RNA, color=percent.mito)) + 
geom_point() + 
scale_colour_gradient(low = "gray90", high = "black") +
stat_smooth(method=lm) +
scale_x_log10() +
scale_y_log10() + 
theme_classic() +
geom_vline(xintercept = 500) +
geom_hline(yintercept = 300) +
facet_wrap(~orig.ident) + 
ggtitle("Postfilter")

ggarrange(prefilter_qc_mito,postfilter_qc_mito,ncol=1,nrow=2,widths = 25,heights = 30)
```

<a  id="Ribosomal_Percentage"></a> 
### Ribosomal Percentage 
```{r Ribo Percentage Plot}
# ribo
prefilter_qc_ribo <- metadata_prefilter %>% ggplot(aes(x=nCount_RNA, y=nFeature_RNA, color=percent.ribo)) + 
geom_point() + 
scale_colour_gradient(low = "gray90", high = "black") +
stat_smooth(method=lm) +
scale_x_log10() +
scale_y_log10() + 
theme_classic() +
geom_vline(xintercept = 500) +
geom_hline(yintercept = 300) +
facet_wrap(~orig.ident) +
ggtitle("Prefilter")

postfilter_qc_ribo <- metadata_postfilter %>% ggplot(aes(x=nCount_RNA, y=nFeature_RNA, color=percent.ribo)) + 
geom_point() + 
scale_colour_gradient(low = "gray90", high = "black") +
stat_smooth(method=lm) +
scale_x_log10() +
scale_y_log10() + 
theme_classic() +
geom_vline(xintercept = 500) +
geom_hline(yintercept = 300) +
facet_wrap(~orig.ident) +
ggtitle("Postfilter")

ggarrange(prefilter_qc_ribo,postfilter_qc_ribo,ncol=1,nrow=2,widths = 10,heights = 8)
```
<a  id="UMIs_transcripts_per_cell"></a> 
### UMIs/transcripts per cell
```{r Visualize the number UMIs/transcripts per cell}
UMIs_transcripts_per_cell_prefilter <- metadata_prefilter %>% 
  	ggplot(aes(color=orig.ident, x=nCount_RNA, fill= orig.ident)) + 
  	geom_density(alpha = 0.2) + 
  	scale_x_log10() + 
  	theme_classic() +
  	ylab("Cell density") +
  	geom_vline(xintercept = 500) + 
    ggtitle("Prefilter")

UMIs_transcripts_per_cell_postfilter <- metadata_postfilter %>% 
  	ggplot(aes(color=orig.ident, x=nCount_RNA, fill= orig.ident)) + 
  	geom_density(alpha = 0.2) + 
  	scale_x_log10() + 
  	theme_classic() +
  	ylab("Cell density") +
  	geom_vline(xintercept = 500) + 
    ggtitle("Postfilter")

ggarrange(UMIs_transcripts_per_cell_prefilter,UMIs_transcripts_per_cell_postfilter,ncol=1,nrow=2,widths = 10,heights = 8)
```
<a  id="Genes Per Cell"></a> 
### genes per cell 
```{r Genes Per Cell}
genes_per_cell_pre <- metadata_prefilter %>% 
  	ggplot(aes(color=orig.ident, x=nCount_RNA, fill=orig.ident )) + 
  	geom_density(alpha = 0.2) + 
  	theme_classic() +
  	scale_x_log10() + 
  	geom_vline(xintercept = 300) + 
    ggtitle("Prefilter")

genes_per_cell_post <- metadata_postfilter %>% 
  	ggplot(aes(color=orig.ident, x=nCount_RNA, fill=orig.ident )) + 
  	geom_density(alpha = 0.2) + 
  	theme_classic() +
  	scale_x_log10() + 
  	geom_vline(xintercept = 300) + 
    ggtitle("Postfilter")

ggarrange(genes_per_cell_pre,genes_per_cell_post,ncol=1,nrow=2,widths = 10,heights = 8)
```
  
```{r }
if (length(samples)>=3){
# Integration
# normalize and identify variable features for each dataset independently
scrna.list <- SplitObject(so_postfilter, split.by = "orig.ident")
scrna.list <- lapply(X = scrna.list, FUN = function(x) {
    x <- NormalizeData(x)
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
})
# select features that are repeatedly variable across datasets for integration
features <- SelectIntegrationFeatures(object.list = scrna.list)

# Perform integration
scrna.anchors <- FindIntegrationAnchors(object.list = scrna.list, anchor.features = features)
scrna.combined <- IntegrateData(anchorset = scrna.anchors)
rm(scrna.anchors)

# Perform an integrated analysis
DefaultAssay(scrna.combined) <- "integrated"
scrna.combined <- ScaleData(scrna.combined, verbose = FALSE)
scrna.combined <- RunPCA(scrna.combined, npcs = 15, verbose = FALSE)
} else{
# Preprocess workflow
# normalize and identify variable features 
so_postfilter <- NormalizeData(so_postfilter)
so_postfilter <- FindVariableFeatures(so_postfilter, selection.method = "vst", nfeatures = 2000)
so_postfilter <- ScaleData(so_postfilter, verbose = FALSE)
so_postfilter <- RunPCA(so_postfilter, npcs = 15, verbose = FALSE)
}

           
```
<a  id="elbow_plot"></a> 
### Linear Dimensionality Reduction 
```{r rna_Elbow_Plot}
if (length(samples)>=3){
    p1 <- ElbowPlot(scrna.combined) + ggtitle("Integrated") + theme(aspect.ratio=5/10) + theme(plot.margin = unit(c(6, 6, 6, 6), "cm"))
    print(p1 + coord_fixed())
} else{
    p1 <- ElbowPlot(so_postfilter) + ggtitle("Elbow Plot") + theme(aspect.ratio=5/10) + theme(plot.margin = unit(c(6, 6, 6, 6), "cm"))
    print(p1 + coord_fixed())
}

```
<a  id="UMAP_Resolutions"></a> 
### UMAP - Resolutions
```{r rna_UMAP_Resolutions_Plot}
# Continue on analysis
if (length(samples)>=3){
    # Continue on integrated analysis
scrna.combined <- RunUMAP(scrna.combined, reduction = "pca", dims = 1:8,verbose=FALSE)
scrna.combined <- FindNeighbors(scrna.combined, reduction = "pca", dims = 1:8,verbose=FALSE)
scrna.combined <- FindClusters(scrna.combined, resolution = c(0.5,0.7,1,1.2,1.4,2),verbose=FALSE)
res_0.5 <- DimPlot(scrna.combined, group.by = "integrated_snn_res.0.5", label = TRUE)
pdf(paste0(out_dir, "res_0.5.pdf"))
dev.off()
res_0.7 <- DimPlot(scrna.combined, group.by = "integrated_snn_res.0.7", label = TRUE)
pdf(paste0(out_dir, "res_0.7.pdf"))
dev.off()
res_1 <- DimPlot(scrna.combined, group.by = "integrated_snn_res.1", label = TRUE)
pdf(paste0(out_dir, "res_1.pdf"))
dev.off()
res_1.2 <- DimPlot(scrna.combined, group.by = "integrated_snn_res.1.2", label = TRUE)
pdf(paste0(out_dir, "res_1.2.pdf"))
dev.off()
res_1.4 <- DimPlot(scrna.combined, group.by = "integrated_snn_res.1.4", label = TRUE)
pdf(paste0(out_dir, "res_1.4.pdf"))
dev.off()
res_2 <- DimPlot(scrna.combined, group.by = "integrated_snn_res.2", label = TRUE)
pdf(paste0(out_dir, "res_2.pdf"))
dev.off()
Idents(scrna.combined) <- resolution
rm(scrna.list)
saveRDS(scrna.combined, paste0(outdir, "scrna.combined_postfilter.seurat.", projectName, ".rds"))
} else {
so_postfilter <- RunUMAP(so_postfilter, reduction = "pca", dims = 1:8,verbose=FALSE)
so_postfilter <- FindNeighbors(so_postfilter, reduction = "pca", dims = 1:8,verbose=FALSE)
so_postfilter <- FindClusters(so_postfilter, resolution = c(0.5,0.7,1,1.2,1.4),verbose=FALSE)
res_0.5 <- DimPlot(so_postfilter, group.by = "RNA_snn_res.0.5", label = TRUE)
pdf(paste0(out_dir, "res_0.5.pdf"))
res_0.5
dev.off()
pdf(paste0(out_dir, "res_0.7.pdf"))
res_0.7 <- DimPlot(so_postfilter, group.by = "RNA_snn_res.0.7", label = TRUE)
res_0.7
dev.off()
pdf(paste0(out_dir, "res_1.pdf"))
res_1 <- DimPlot(so_postfilter, group.by = "RNA_snn_res.1", label = TRUE)
dev.off()
pdf(paste0(out_dir, "res_1.2.pdf"))
res_1.2 <- DimPlot(so_postfilter, group.by = "RNA_snn_res.1.2", label = TRUE)
dev.off()
pdf(paste0(out_dir, "res_1.4.pdf"))
res_1.4 <- DimPlot(so_postfilter, group.by = "RNA_snn_res.1.4", label = TRUE)
dev.off()
Idents(so_postfilter) <- resolution
saveRDS(so_postfilter, paste0(outdir, "scrna_postfilter.seurat.", projectName, ".rds"))
print(res_0.5)
print(res_0.7)
print(res_1)
print(res_1.2)
print(res_1.4)
}
```

<a  id="UMAP"></a> 
```{r rna_UMAP}
### UMAP - Treatment
if (length(samples)>=3){
p1 <- DimPlot(scrna.combined, reduction = "umap", group.by = "orig.ident")
print(p1 + coord_fixed())
#p2 <- DimPlot(scrna.combined, reduction = "umap", label = TRUE, repel = TRUE)
#print(p2 + coord_fixed())
p3 <- DimPlot(scrna.combined, reduction = "umap", split.by = "orig.ident", ncol = 4)
print(p3 + coord_fixed())
p4 <- DimPlot(scrna.combined, reduction = "umap", group.by = "treatment")
print(p4 + coord_fixed())
rm(p1)
#rm(p2)
rm(p3)
rm(p4)
} else {
p1 <- DimPlot(so_postfilter, reduction = "umap", group.by = "orig.ident")
print(p1 + coord_fixed())
#p2 <- DimPlot(so_postfilter, reduction = "umap", label = TRUE, repel = TRUE)
#print(p2 + coord_fixed())
p3 <- DimPlot(so_postfilter, reduction = "umap", split.by = "orig.ident", ncol = 4)
print(p3 + coord_fixed())
p4 <- DimPlot(so_postfilter, reduction = "umap", group.by = "treatment")
print(p4 + coord_fixed())
rm(p1)
#rm(p2)
rm(p3)
rm(p4)
}
```

```{r }
if (length(samples)>=3){
DefaultAssay(scrna.combined) <- "RNA"
all.genes <- rownames(scrna.combined)
scrna.combined <- ScaleData(scrna.combined,features = all.genes,verbose = FALSE)
#nk.markers <- FindConservedMarkers(scrna.combined, ident.1 = 0, grouping.var = "orig.ident", verbose = FALSE)
# FindAllMarkers
scrna.markers <- FindAllMarkers(scrna.combined, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25,verbose=FALSE,test.use = "MAST")
names(scrna.markers)[names(scrna.markers) == "gene"] <- "geneSymbol"
scrna.markers$geneID <- geneTable$geneID[match(scrna.markers$geneSymbol, geneTable$geneSymbol)]
write.table(scrna.markers, paste0(out_dir, "FindAllMarkers.clusters.xls"), sep = "\t", row.names = F)
} else {
# Step 5: Identify conserved cell type markers
# For performing differential expression after integration, we switch back to the original data
# Identify conserved cell type markers
#nk.markers <- FindConservedMarkers(so_postfilter, ident.1 = 0, grouping.var = "orig.ident")
scrna.markers <- FindAllMarkers(so_postfilter, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25,verbose=FALSE,test.use = "MAST")
names(scrna.markers)[names(scrna.markers) == "gene"] <- "geneSymbol"
write.table(scrna.markers, paste0(out_dir, "FindAllMarkers.clusters.xls"), sep = "\t", row.names = F)
}
```
                                                        
```{r }
# top 10
top10<- scrna.markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_log2FC)
write.table(top10, paste0(out_dir, "FindAllMarkers.clusters.top10.xls"), sep = "\t", col.names = NA)

# top 25
top25 <- scrna.markers %>% group_by(cluster) %>% top_n(n = 25, wt = avg_log2FC)
write.table(top25, paste0(out_dir, "FindAllMarkers.clusters.top25.xls"), sep = "\t", col.names = NA)

# top 50
top50 <- scrna.markers %>% group_by(cluster) %>% top_n(n = 50, wt = avg_log2FC)
write.table(top50, paste0(out_dir, "FindAllMarkers.clusters.top50.xls"), sep = "\t", col.names = NA)

# Step 6: Top 3 identified genes, feature plot, dotplot
topN <- Extract_Top_Markers(scrna.markers, num_genes = 25, named_vector = FALSE, make_unique = TRUE, gene_column = "geneSymbol")
```

<a  id="mgt"></a> 
```{r Marker_Gene_Table}
# Create Interactive Table
as_tibble(unique(scrna.markers)) %>% arrange(p_val_adj) -> scrna.markers
scrna.markers = scrna.markers %>% dplyr::select(cluster, geneSymbol, p_val, avg_log2FC,
                                             p_val_adj,pct.1,pct.2)
final_df <- data.frame(           "cluster"     = scrna.markers$cluster,
                                  "gene"       = scrna.markers$geneSymbol,
                                  "p_val"      = scrna.markers$p_val,
                                  "avg_log2FC" = scrna.markers$avg_log2FC,
                                  "p_val_adj"  = scrna.markers$p_val_adj,
                                  "pct.1"      = scrna.markers$pct.1,
                                  "pct.2"      = scrna.markers$pct.2)            
DT::datatable(final_df,class = 'cell-border stripe',rownames=F,filter='top',
              editable = TRUE, extensions = 'Buttons', options = list(
                dom = 'Bfrtip',
                buttons = c('copy','csv','excel','pdf','print')
              ),escape = FALSE)
            
```
<a  id="fp"></a> 
```{r Feature_Plot}
if (exists("scrna.combined")==TRUE){ so_postfilter<-scrna.combined }
# Feature plot
pdf(paste0(out_dir, "combined.top25markers.pdf"))
ggp = list()
for (marker in topN){
    ggp[[marker]]=FeaturePlot(so_postfilter, features=marker)
    print(ggp[[marker]])
}
dev.off()

pdf(paste0(out_dir, "combined.top25markers.geneSymbol.pdf"))
ggp = list()
for (marker in topN){
    ggp[[marker]]=FeaturePlot(so_postfilter, features=marker) + ggtitle(geneTable$geneSymbol[match(marker, geneTable$geneID)])
    print(ggp[[marker]])
}
dev.off()


#DoHeatmap(scrna.combined, features = topN$geneID, size = 2, draw.lines = T, angle = 45, hjust = 0.2) + theme(axis.text.y = element_text(size = 5)) + NoLegend() + scale_y_discrete(breaks=topN$geneID,
#        labels=geneTable$geneSymbol[match(topN$geneID, geneTable$geneID)])
#ggsave(paste0(outdir, "top25markers.heatmap.integrated.geneSymbol.pdf"), width = 8, height = 12)
#  
```

```{r Customer Markers, Feature Plot, Dot Plot}
remove_markers <- setdiff(markers,row.names(so_postfilter))
markers <- markers[!markers%in%remove_markers]
pdf(paste0(out_dir, "combined.markers.geneSymbol.pdf"))
ggp = list()
for (marker in markers){
    ggp[[marker]]=FeaturePlot(so_postfilter, features=marker) + ggtitle(marker)
    print(ggp[[marker]])
}
dev.off()

```
 ---  

scRNA seq report well module v.`r {scrna_seq_sample_module_version}`, Brigham and Women's Bioinformatics and Genomics Hub  

```{r scrna_cleanup_cluster_annotation_, include = FALSE}
module_vars <- setdiff(ls(), orig_workspace)
rm(list=module_vars)
gc()
```
